---
title: "604 Final Project"
author: "Nandni Talreja"
date: "04/03/2024"
format:
  html:
    df-print: paged
    css: "styles.css"
    embed-resources: true
    self-contained-math: true
categories:
  - Final Project
---

# Navigating the Cyber Highway: Assessing the knowledge and Awareness of Vehicle Cybersecurity among Law Enforcement Officers (LEOs)

## Details of the Final Project
The primary goal of the written project (as well as the recorded slideshow presentation) is to apply techniques and principles covered in the readings, homework, and lectures. Consider these materials as your rubric! That is, your grade will be based primarily on the quality of communication, in keeping with the content of the course. 
Project Dataset - Law Cybersecurity Survey (unpublished, conducted by Nandni Talreja in the Human Performance Lab under Professor Shannon Roberts)
Expected Audience - Police Officers and Police Stations (for training)

## Project Propoal
I plan to study the survey data collected in a study of Law Enforcement Officers (LEOs) awareness of vehicle cybersecurity. The study was conducted by me in the Human Performance Lab under Professor Shannon Roberts. The aim of the study was to investigate the association between cybersecurity training and officers' understanding of cybersecurity concepts applicable to police vehicle systems. By taking into account demographic variables such as age groups, education levels, years of experience, and hours spent in police vehicles, I plan to study how these variables interact with the level of cybersecurity training to influence officers' understanding of cybersecurity in police vehicles.
I plan to conduct statistical analysis to understand the connection between the demographic variables and the understanding of vehicular cybersecurity. I plan to use regression modeling and hypothesis testing to determine the level of understanding LEOs had of vehicular cybersecurity before and after their training. I also plan to look into the difference in understanding in the two groups - Training and Sham. The aim of my study is to contribute to the understanding of cybersecurity training, both vehicular and otherwise, for LEOs. The aim is to understand if the training model created works best for enhancing officers' cybersecurity awareness specific to police vehicle systems. With my study, I hope to provide strategies for improving the training models relating to cybersecurity in the Law Enforcement Department.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries

library(ggplot2)
library(dplyr)
library(ltm)
library(psych)
library(tidyverse)
library(readxl)
library(haven) #for loading other datafiles (SAS, STATA, SPSS, etc.)
library(stringr) # if you have not installed this package, please install it.
library(lubridate)
library(knitr)
library(forcats)

```

### Importing the dataset and cleaning it up to help with analysis

```{r}

# original dataset from the study
PoliceData <- read.csv("dataset", header = TRUE)
head(PoliceData)

# cleaning and mutating the data

# #removing the first two rows
PoliceData <- PoliceData[-c(1,2),]
PoliceData <- PoliceData[-c(1),]

#removing our test responses
PoliceData <- PoliceData[which.max(PoliceData$IPAddress == "170.63.193.171") : nrow(PoliceData), ]

#mutating Education
PoliceData <- PoliceData %>%
  mutate(Q7 = case_when(
    Q7 == "Doctorate degree (for example: PhD, EdD)" ~ "Doctorate",
    Q7 == "Associates degree (for example: AA, AS)" ~ "Associates",
    Q7 == "Bachelors degree (for example: BA, BS)" ~ "Bachelors",
    Q7 == "Masters degree (for example: MA, MS, MEng, MEd, MSW, MBA)" ~ "Masters",
    TRUE ~ Q7  
  ))

# changing "14 years" to "14" in the years of experience column
PoliceData <- PoliceData %>%
  mutate(Q6 = as.character(Q6)) %>%
  mutate(Q6 = case_when(Q6 == "14 years" ~ "14",
                        TRUE ~ Q6))

# changing 20-Oct back to 10-20 hours
PoliceData <- PoliceData %>%
  mutate(Q8 = as.character(Q8)) %>%
  mutate(Q8 = case_when(Q8 == "20-Oct" ~ "10 -20",
                        TRUE ~ Q8))

head(PoliceData) # just for my sanity

colnames(PoliceData) # to understand what columns I want to add to the new dataset that I will use for this project

# creating a new dataset with just the columns that I am going to be using
data <- PoliceData[, c("Q32", "Q1", "Q2", "Q6", "Q7", "Q8", "Q9", "Q12", "Q13", "Q14", "Q15", "Q16", "Q18", "Q19", "Q22", "Q23", "Q25", "Q28")]

# mutating the years column to be in groups for easy visualization
data <- data %>%
  mutate(Q6 = as.numeric(Q6)) %>%
  mutate(years = case_when(
    Q6 < 5 ~ "less than 5 years",
    Q6 >= 5 & Q6 < 10 ~ "5-10",
    Q6 >= 10 & Q6 < 15 ~ "10-15",
    Q6 >= 15 & Q6 <20 ~ "15-20",
    Q6 >= 20 & Q6 <25 ~ "20-25",
    Q6 >= 25 & Q6 <30 ~ "25-30"
  ))

head(data) # sanity check

```

## Basic Statistics and Summary of the Data
```{r}

#average duration
mean(as.numeric(PoliceData$Duration..in.seconds.))

#changing age to numeric
data$Q1 <- as.numeric(data$Q1)
# summary
summary(data)

# tables
# gender
gendert <- table(data$Q2)
gt <- as.data.frame(gendert)
kable(gt, caption = "Gender")

# years of experience
yearst <- table(data$Q6)
yt <- as.data.frame(yearst)
kable(yt, caption = "Years of Experience")

yeart <- table(data$years)
yyt <- as.data.frame(yeart)
kable(yyt, caption = "Years of Experience")

# degree
degreet <- table(data$Q7)
dt <- as.data.frame(degreet)
kable(dt, caption = "Highest Education Level")

# hours in police vehicle
hourst <- table(data$Q8)
ht <- as.data.frame(hourst)
kable(ht, caption = "Hours in Police Vehicle")

# cyber cases
cybercasest <- table(data$Q12)
cct <- as.data.frame(cybercasest)
kable(cct, caption = "Cyber Cases in the Agency")

# cyber training
cybertraint <- table(data$Q14)
ctt <- as.data.frame(cybertraint)
kable(ctt, caption = "Cyber Training in the Agency")

militaryt <- table(data$Q9)
mt <- as.data.frame(militaryt)
kable(mt, caption = "Military Experience")

cyberunitt <- table(data$Q13)
cut <- as.data.frame(cyberunitt)
kable(cut, caption = "Cyber Unit in the Agency")

caseworkt <- table(data$Q15)
cwt <- as.data.frame(caseworkt)
kable(cwt, caption = "Cases Worked")

vcybert <- table(data$Q16)
vct <- as.data.frame(vcybert)
kable(vct, caption = "Experience with Vehicle CyberAttacks")

cybervehiclet <- table(data$Q18)
cvt <- as.data.frame(cybervehiclet)
kable(cvt, caption = "Police Vehicle Susceptibility of CyberAttacks")

concernt <- table(data$Q19)
cont <- as.data.frame(concernt)
kable(cont, caption = "Vehicle Cyberattacks Concern")


```

## Basic Demographic Plots and Analysis
``` {r Numeric Questions, echo = FALSE}
# This section plots the questions in the police survey

#changing years to numeric
data$Q6 <- as.numeric(data$Q6)

#Gender
ggplot(data = data, aes(x = Q2)) + 
  stat_count(fill = "lightgreen", color = "black") + 
  xlab("Gender")

#Years of experience
ggplot(data = data, aes(x = Q6)) + 
  geom_bar(fill = "yellow", color = "black") + 
  xlab("Years of Experience")

ggplot(data = data, aes(x = years)) + 
  geom_bar(fill = "yellow", color = "black") + 
  xlab("Years of Experience")

#Education
ggplot(data = data, aes(x = Q7)) + 
  geom_bar(stat = "count", fill = "lavender", color = "black") + 
  xlab("Education")

#military experience
ggplot(data = data, aes(x = Q9)) + 
  geom_bar(stat = "count", fill = "brown", color = "black") + 
  xlab("Military Experience")

#cybercrime unit in police agency
ggplot(data = data, aes(x = Q13)) + 
  geom_bar(stat = "count", fill = "red", color = "black") + 
  xlab("CyberCrime unit in Police Agency")

#cybercrime cases handline in police agency
ggplot(data = data, aes(x = Q12)) + 
  geom_bar(stat = "count", fill = "green", color = "black") + 
  xlab("CyberCrime cases in Police Agency")

#police agency training for cybercrimes
ggplot(data = data, aes(x = Q14)) + 
  geom_bar(stat = "count", fill = "yellow", color = "black") + 
  xlab("CyberCrime Training in Police Agency")

#Time spent in car
ggplot(data = data, aes(x = Q8)) + 
  geom_bar(stat = "count", fill = "pink", color = "black") + 
  xlab("Hours spent in car")

#police vehicle susceptibility of cyberattacks
ggplot(data = data, aes(x = Q18)) + 
  geom_bar(stat = "count", fill = "white", color = "black") + 
  xlab("Police Vehicle Susceptibility of CyberAttacks")



```

## Now, adding in the "Experimental" and "Sham" aspect to the data
```{r}
# {r}
# Dividing participants into Experimental and Sham groups
# Experimental Participants - 20133, 76685, 40329, 55596, 69301, 97268, 13457
# Sham Participants - 26347, 30034, 42937, 37144, 59271, 47969, 88218

data <- data %>%
  mutate(stgroup = ifelse(Q32 %in% c("20133", "76685", "40329", "55596", "69301",  "97268", "13457"), "Experimental", "Sham"))

head(data) # sanity check

# visualizing the experimental and sham groups to make sure there are equal number of people in both groups
ggplot(data = data, aes(x = stgroup)) + 
  geom_bar(stat = "count", fill = "blue", color = "black") + 
  xlab("Distribution of Participants in the study")

```

## Plotting the variables with "Experimental" and "Sham" groups as the independent variable
```{r}

training_19 <- data %>%
  count(stgroup, Q19) %>%
  ggplot(aes(x = Q19, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Are cyber attacks on police vehicles a serious concern for you?") +
  theme_minimal()
training_19

training_18 <- data %>%
  count(stgroup, Q18) %>%
  ggplot(aes(x = Q18, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  xlab("Police Vehicle Susceptibility of CyberAttacks") +
  theme_minimal()
training_18

training_13 <- data %>%
  count(stgroup, Q13) %>%
  ggplot(aes(x = Q13, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("CyberCrime unit in Police Agency") +
  theme_minimal()
training_13

training_12 <- data %>%
  count(stgroup, Q12) %>%
  ggplot(aes(x = Q12, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("CyberCrime cases in Police Agency") +
  theme_minimal()
training_12

training_14 <- data %>%
  count(stgroup, Q14) %>%
  ggplot(aes(x = Q14, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("CyberCrime Training in Police Agency") +
  theme_minimal()
training_14

training_8 <- data %>%
  count(stgroup, Q8) %>%
  ggplot(aes(x = Q8, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Hours spent in car") +
  theme_minimal()
training_8

training_years <- data %>%
  count(stgroup, years) %>%
  ggplot(aes(x = years, y = n, fill = stgroup)) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Hours spent in car") +
  theme_minimal()
training_years

```

# Correlation tests
``` {r}

# making a subset with certain questions and doing a correlation test on it
data <- data %>%
  mutate(gender = case_when(
    Q2 == "Male" ~ "0",
    Q2 == "Female" ~ "1",
    TRUE ~ Q2  
  ))
data$gender <- as.numeric(data$gender)

unique(data$Q7)

data <- data %>%
  mutate(degree = case_when(
    Q7 == "Some College" ~ "0",
    Q7 == "Associates" ~ "1",
    Q7 == "Bachelors" ~ "2",
    Q7 == "Masters" ~ "3",
    Q7 == "Doctorate" ~ "4",
    Q7 != "Some College" & Q7 != "Associates" & Q7 != "Bachelors" & Q7 != "Masters" & Q7 != "Doctorate" ~ as.character(Q7)
  ))
data$degree <- as.numeric(data$degree)

data <- data %>%
  mutate(time = case_when(
    Q8 == "10 -20" ~ "0",
    Q8 == "20-30" ~ "1",
    Q8 == "30-40" ~ "2",
    Q8 == "40+" ~ "3",
    Q8 != "10 -20" & Q8 != "20-30" & Q8 != "30-40" & Q8 != "40+" ~ as.character(Q8)
  ))
data$time <- as.numeric(data$time)

#creating a dataset with just these 8 questions
adata <- data[,c("Q1", "gender", "Q6", "time", "Q12", "Q13", "Q14", "Q15", "Q18", "Q19")]
adata[,5:10]<-ifelse(adata[,5:10]=="Yes",1,0)

#correlation test
cor(adata) 

#calculating Cronbach's alpha for these 8 questions
#alpha(x=adata[,1:4], check.keys = TRUE)
#alpha(x=adata[,5:7], check.keys = TRUE)

```

# Regression Analysis
```{r}
# let's conduct some regression models now
# first we're going to change the stgroup values to numeric - so 1 for experimental and 0 for sham
data <- data %>%
  mutate(groupnum = case_when(
    stgroup == "Experimental" ~ 1,
    stgroup == "Sham" ~ 0
  ))
head(data)

data$groupnum = as.numeric(data$groupnum)
data$groupnum = as.factor(data$groupnum)

data$stgroup <- as.factor(data$stgroup)
data$Q19 <- as.factor(data$Q19)
data$Q16 <- as.factor(data$Q25)

# let's compare the stgroup with understand concerns of vehicle cybersecurity
summary(glm(Q19 ~ stgroup, data = data, family = binomial))

summary(glm(Q25 ~ stgroup, data = data, family = binomial))

```




